## Experimental Design {-}

### Completely Random Design {-}

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

- Treatment Structure: 2 x 3 Factorial Treatment, both Fixed

- Model: $y_{ijk} = \mu + \alpha_i + \beta_j + \alpha \beta_{ij} + e_{ijk}$
- Treatments: $\alpha_i = \text{supp, } \beta_j = \text{dose}$
- Fixed Effects: $\alpha_1 = \beta_1 = \alpha \beta_{1j} = \alpha \beta_{i1} = 0$
- Random Effects: $e_{ijk} = N(0, \sigma^2_e)$

```{r a13, comment=NA, fig.width=8, fig.height=5, warning=FALSE, message=FALSE}
library(lsmeans)
library(reshape2)
library(car)
library(plyr)

data("ToothGrowth")

## Data is numeric, but we need to force it to be a factor for the model
ToothGrowth$dose.factor = as.factor(ToothGrowth$dose)
summary(ToothGrowth)

## Even observations per treatment group
table(ToothGrowth$supp, ToothGrowth$dose)

## ANOVA model
mdl = lm(len ~ supp * dose.factor, data = ToothGrowth)
anova(mdl)

## Least Squares Means
lsmeans(mdl, specs = c("supp", "dose.factor"))

## Differences in Means
TukeyHSD(aov(mdl))

```

Are the necessary conditions for hypothesis testing present?

- Normality: Residuals appear normally distributed per the residual normal reference plot and shapiro-wilks test
- Equal Variance: Brown-Forsythe-Levene test and residual plot supports equal variance
- Independence: No correlation in the residuals per the Durbin Watson test and plots of variables against residuals

Conditions for hypothesis testing appears to be satisfied

```{r a14, comment=NA, fig.width=8, fig.height=5, warning=FALSE, message=FALSE}

## Normality of Residuals
shapiro.test(mdl$residuals); qqnorm(mdl$residuals); qqline(mdl$residuals)

## Equal Variance of Residuals
leveneTest(mdl)
plot(x = mdl$fitted.values, y = (mdl$residual - mean(mdl$residuals))/sd(mdl$residuals),
     xlab = "Fitted Values", ylab = "Standardized Residuals",
     main = "Equal Variance Residual Plot")

## Independence of Residuals
par(mfrow = c(1, 2))
plot(ToothGrowth$dose, mdl$residuals, xlab = "Dose", ylab = "Residuals")
plot(ToothGrowth$supp, mdl$residuals, xlab = "Supp", ylab = "Residuals")

## Test for correlation in the residuals
durbinWatsonTest(mdl)

```

- Group doses so that each dose is not statistically different than any other dose in the group:
    + The interaction between dose and supp are significant so we need to assess the differences in dose per each level of supp.
    + OJ: {.5}, {1, 2}
    + VC: {.5}, {1}, {2}
- Group supps so that each supp is not statistically different than any other supp in the group:
    + The interaction is significant so we need to assess the supps at each level of dose
    + .5: {OJ}, {VC}
    + 1: {OJ}, {VC}
    + 2: {OJ, VC}


```{r a15, comment=NA, fig.width=8, fig.height=5, echo=FALSE}

## Calculate the mean and sd for each combination of supp and dose
x = ddply(ToothGrowth, .(supp, dose), summarize, mean = mean(len), sd = sd(len))

## add 95% confidence intervals
x$lwr = with(x, mean - qt(p = .975, df = 9) * sd/sqrt(10))
x$upr = with(x, mean + qt(p = .975, df = 9) * sd/sqrt(10))

x1 = subset(x, supp == "OJ")
x2 = subset(x, supp == "VC")

## Make a blank plot
plot(x = ToothGrowth$dose,
     y = ToothGrowth$len,
     xlab = "Dose",
     ylab = "Tooth Length",
     type = "n",
     ylim = c(6, 29),
     main = "Interaction Plot")

## Add points for the means
points(x = x1$dose, y = x1$mean, pch = 3, lwd = 4, col = "blue", cex = 1.1)
points(x = x2$dose, y = x2$mean, pch = 4, lwd = 4, col = "red", cex = 1.1)

## Add lines for interaction
lines(x = x1$dose, y = x1$mean, lty = 2)
lines(x = x2$dose, y = x2$mean, lty = 2)

## Add segments for confidence intervals
for (i in 1:6) {
  segments(x0 = x$dose[i], y0 = x$lwr[i], y1 = x$upr[i])
  segments(x0 = x$dose[i]-.05, x1 = x$dose[i]+.05, y0 = x$lwr[i])
  segments(x0 = x$dose[i]-.05, x1 = x$dose[i]+.05, y0 = x$upr[i])
}

## Add legend
legend("bottomright", c("Orange Juice", "Vitamin C"),
       pch = c(3, 4), bty = "n", cex = 1.1, col = c("blue", "red"))

```


### Random Complete Block Design {-}

An experiment was conducted to compare four different pre-planting treatments for soybeen seeds. A fifth treatment, consisting of not treating the seeds was used as a control. The experimental area consisted of four fields. There are notable differences in the fields. Each field was divided into five plots and one of the treatments was randomly assigned to a plot within each field.

- Treatment Structure: 1 Single Treatment with 5 levels
- Response: The number of plants that failed to emerge out of 100 seeds planted per plot.

- Model: $y_{ij} = \mu + \alpha_i + \beta_j + e_{ij}$
- Treatments: $\alpha_i = \text{Seed, } \beta_j = \text{Field, } \alpha_5 = \text{Control}$
- Fixed Effects: $\alpha_5 = \beta_1 = 0$
- Random Effects: $e_{ij} = N(0, \sigma^2_e)$

---

```{r a16, echo=FALSE, results='asis'}
library(knitr)
library(reshape2)
library(pander)

soy = data.frame(
  Treatment = c("Avasan", "Spergon", "Semaesan", "Fermate", "Control"),
  Field.1 = c(2, 4, 3, 9, 8),
  Field.2 = c(5, 10, 6, 3, 11),
  Field.3 = c(7, 9, 9, 5, 12),
  Field.4 = c(11, 8, 10, 5, 13)
)

pandoc.table(soy, caption = "Soy Bean Data")

soy = melt(soy, id.vars = "Treatment", variable.name = "Field", value.name = "Count")

```


```{r a17, comment=NA}
## Make the control treatment the default level
soy$Treatment = relevel(soy$Treatment, ref = "Control")

## We only have one rep per treatment so there are not enough DF to measure the interaction
mdl = lm(Count ~ Field + Treatment, data = soy)
anova(mdl)

```

---

#### Comparison of means vs control

Since we have a control variable we want to know if any of the treatment means are significantly lower than the control mean.

```{r a18, comment=NA, message=FALSE, warning=FALSE}
library(multcomp)

Dunnet = glht(mdl, linfct = mcp(Treatment = "Dunnet"), alternative = "less")
summary(Dunnet)

```

We have significant evidence that only Avasan and Fermate are significantly lower than the control. Are they significantly different from each other?

```{r a19, comment=NA, message=FALSE, warning=FALSE}

TukeyHSD(aov(mdl))

```

There is not significant evidence between the difference in means between any of the treatments.
