---
output: html_document
---

## **Random Complete Block Design**

---

#### **Experimental Design**

An experiment was conducted to compare four different pre-planting treatments for soybeen seeds. A fifth treatment, consisting of not treating the seeds was used as a control. The experimental area consisted of four fields. There are notable differences in the fields. Each field was divided into five plots and one of the treatments was randomly assigned to a plot within each field.

- **Randomization**: RCBD  
- **Treatment Structure**: 1 Single Treatment with 5 levels
- **Response**: The number of plants that failed to emerge out of 100 seeds planted per plot.

---

#### **Statistical Model**
- **Model**: $y_{ij} = \mu + \alpha_i + \beta_j + e_{ij}$  
- **Parameters**: $\alpha_i = \text{Seed, } \beta_j = \text{Field, } \alpha_5 = \text{Control}$  
- **Fixed Effects**: $\alpha_5 = \beta_1 = 0$
- **Random Effects**: $e_{ij} = N(0, \sigma^2_e)$  

---

```{r a, echo=FALSE, results='asis'}
library(knitr)
library(reshape2)
library(pander)

soy = data.frame(
  Treatment = c("Avasan", "Spergon", "Semaesan", "Fermate", "Control"),
  Field.1 = c(2, 4, 3, 9, 8),
  Field.2 = c(5, 10, 6, 3, 11),
  Field.3 = c(7, 9, 9, 5, 12),
  Field.4 = c(11, 8, 10, 5, 13)
)

pandoc.table(soy, caption = "Soy Bean Data")

soy = melt(soy, id.vars = "Treatment", variable.name = "Field", value.name = "Count")

```


```{r b, comment=NA}
## Make the control treatment the default level
soy$Treatment = relevel(soy$Treatment, ref = "Control")

## We only have one rep per treatment so there are not enough DF to measure the interaction
mdl = lm(Count ~ Field + Treatment, data = soy)
anova(mdl)

```

---

#### Comparison of means vs control

Since we have a control variable we want to know if any of the treatment means are significantly lower than the control mean.

```{r c, comment=NA, message=FALSE, warning=FALSE}
library(multcomp)

Dunnet = glht(mdl, linfct = mcp(Treatment = "Dunnet"), alternative = "less")
summary(Dunnet)

```

We have significant evidence that only Avasan and Fermate are significantly lower than the control. Are they significantly different from each other?

```{r d, comment=NA, message=FALSE, warning=FALSE}

TukeyHSD(aov(mdl))

```

There is not significant evidence between the difference in means between any of the treatments.




















