---
output: html_document
---

## **Univariate Distributions and Parameter Estimation**

---

##### **Techniques Used**
  * Various Distribution Plots
  * Formal Tests for Distribution Fit
  * Hypothesis Testing

---

```{r a1, comment=NA, fig.width=8, fig.height=7}
## QQ plot for normal distribution
set.seed(10)
x1 = rnorm(20)

## Use the QQ function for testing normality
par(mfrow = c(2, 2))
qqnorm(x1)
qqline(x1)
boxplot(x1, main = "Boxplot")
hist(x1, freq = FALSE, main = "Histogram with Density Curve")
lines(density(x1))
plot(ecdf(x1), main = "Empiracle CDF")

```

QQ plot indicates the data might be normal by remaining close to the line. The Box plot, histogram, and density curve all support this assumption. 

Formal tests all agree that the data are from the normal distribution. Shapiro Wilk is considered the best for test for testing normality. 

```{r a2, comment=NA, warning=FALSE, message=FALSE}
## Are the data from a normal distribution?
shapiro.test(x1)     ## Shapiro-Wilk Test
library(nortest)
ad.test(x1)          ## Anderson Darling Test
ks.test(x1, 'pnorm') ## Kolmogorov-Smirnoff Test

```

---

#### **Hypothesis Testing**

$$H_0: \mu = 0, H_a: \mu \ne 0$$

```{r a3, comment=NA, warning=FALSE, message=FALSE}
## Is the true mean equal to 0?
t.test(x1)
## What is the power of our test?
power.t.test(n = length(x1), delta = mean(x1), sd = sd(x1), sig.level = .05, type = "one.sample")

```



```{r a4, comment=NA, warning=FALSE, message=FALSE}
x1 = rnorm(100)
## What is the power of our test?
power.t.test(n = length(x1), delta = .25, sd = 1, sig.level = .05, type = "one.sample")


```

