---
output: html_document
---

## **Distribution Fitting**

---

##### **Techniques Used**
  * Assessing Distributions Visually
  * Formal Tests for Distribution Fit
  * Hypothesis Testing
  * Base Plot

---

##### **Normal Distribution**

```{r a1, comment=NA, fig.width=8, fig.height=7}
library(nortest)
library(MASS)

## Draw some random data
set.seed(10)
x1 = rnorm(20)

## Distribution plots
par(mfrow = c(2, 2))
qqnorm(x1)
qqline(x1)
boxplot(x1, main = "Boxplot")
hist(x1, freq = FALSE, main = "Histogram with Density Curve")
lines(density(x1))
plot(ecdf(x1), main = "Empiracle CDF")

```

QQ plot indicates the data might be normal by remaining close to the line. The Box plot, histogram, and density curve all support this assumption. 

Formal tests all agree that the data are from the normal distribution. Shapiro Wilk is considered the best for test for testing normality. 

```{r a2, comment=NA, warning=FALSE, message=FALSE}
## Are the data from a normal distribution?
shapiro.test(x1)     ## Shapiro-Wilk Test
ad.test(x1)          ## Anderson Darling Test
ks.test(x1, 'pnorm') ## Kolmogorov-Smirnoff Test

```

---

##### **Chi-squared Distribution**

```{r a3, comment=NA, warning=FALSE, message=FALSE, fig.width=8, fig.height=7}
## Draw some random data
set.seed(10)
x2 = rchisq(n = 20, 2)

## Estimate the DF parameter by maximum likelihood 
fitdistr(x = x2, dchisq, start = list(df = 2))

## Input the estimate from MLE
ks.test(x = x2, y = pchisq, df = 1.8140625)

## Distribution plots
par(mfrow = c(2, 2))
qqplot(qchisq(ppoints(20), df = 1.8140625), x2)
qqline(x2, distribution = function(p) qchisq(p, df = 1.8140625))
boxplot(x2, main = "Boxplot")
hist(x2, freq = FALSE, main = "Histogram with Density Curve")
lines(density(x2))
plot(ecdf(x2), main = "Empiracle CDF")

```

```{r a4, comment=NA, warning=FALSE, message=FALSE, fig.width=8, fig.height=6}

## Plot the Chi-Squared Distribution with multiple degrees of freedom.
## As DF increases, the distibution converges to the normal distribution
x = seq(.01, 15, by = .05)

plot(x = c(0, 15), y = c(0, .3), type = "n", 
     xlab = "", ylab = "", main = "Chi-Squared Distribution")
lines(x = x, y = dchisq(x, df = 1), col = "red", lwd = 2)
lines(x = x, y = dchisq(x, df = 2), col = "blue", lwd = 2)
lines(x = x, y = dchisq(x, df = 3), col = "green", lwd = 2)
lines(x = x, y = dchisq(x, df = 5), col = "purple", lwd = 2)
lines(x = x, y = dchisq(x, df = 10), col = "orange", lwd = 2)

legend(12, .30,
       c("DF = 1", "DF = 2", "DF = 3", "DF = 5", "DF = 10"),
       lty = rep(1, 5),
       lwd = rep(2, 5),
       col = c("red", "blue", "green", "purple", "orange"))

```


---